We set up data for our paper. We start with

- list of epenthetic stems. 111 verbs.
- 6 suffixes. 

We search the Hungarian Webcorpus using the web interface.

We exclude a set of verbs:

- excluded: defective verbs (18 verbs)
- excluded: ambiguous verbs
- excluded: forms tagged as nouns by the parser (4 forms (_not_ verbs))
- excluded: WITHOUT at least 1 v form AND 1 nv form

We created log odds pairs of v and nv forms.

We created a set of predictor tags.

```{r }
try(setwd("/Users/pracz/Work/Bristol/RaczRebrusTorkenczy2018/"))

library(stringr)
library(stringi)
library(reshape2)
library(plyr)
library(knitr)
library(dplyr)
library(knitr)
# library(arm) # dplyr clashes with arm somehow

# vowelharmony functions
VHround <- function(x){
  y <- ifelse(
    grepl('[öü].$', d$v_base),
      gsub('O', 'ö', x),
        x
  )
  y <- ifelse(
    grepl('e.$', d$v_base),
      gsub('O', 'e', y),
      gsub('O', 'o', y)
  )
return(y)
}

VHfront <- function(x){ x <- ifelse(
  grepl('[aou].$', d$v_base),
    gsub('A','a', x),
      gsub('A','e', x)
)
return(x)
}
```

## 1. Take list of stems, list of suffixes, create query list

```{r }
# take list of stems
d <- read.delim('initial_stem_list.txt', header=F) 
d <- as.character(d$V1)
d <- sub('h.t$','', d)

# remove defective stems (hanyatlik, csuklik)
# d[c(17,18,19,23,27,28,32,33,36,40,53,59,70,84,94,96,97,98)]

d <- d[-c(17,18,19,23,27,28,32,33,36,40,53,59,70,84,94,96,97,98)]

# take list of suffixes
suffixes <- read.delim('initial_suffix_list.txt', sep=',')

# generate all possible forms
d <- merge(d,suffixes)
names(d)[1] <- 'v_base'
d$nv_base <- sub('.(.)$','\\1', d$v_base) 

d$stem <- sub('.$', '', d$nv_base)

d <- d[order(d$stem),c(6,2,1,5,3,4)]
d %>% head(12) %>% kable

d$v_form <- paste(d$v_base,d$v_suffix, sep='')
d$v_form <- VHround(d$v_form)
d$v_form <- VHfront(d$v_form)

d$nv_form <- paste(d$nv_base,d$nv_suffix, sep='')
d$nv_form <- VHround(d$nv_form)
d$nv_form <- VHfront(d$nv_form)

d$exponent <- paste("[", d$v_suffix, "]", sep='')

ql <- c(d$v_form,d$nv_form)

# list of forms for query
# ql %>% head

write.csv(ql, 'initial_query_list.txt', row.names=F)

# list of forms to pair up stems with returns from query
d %>% select(stem, v_form, nv_form) %>% melt(id.vars = 'stem', value.name='word') %>% select(stem,word) -> d2

# list of pairs to long list
d3a <- d[,c('stem','category','exponent','v_base','v_suffix','v_form')]
d3b <- d[,c('stem','category','exponent','nv_base','nv_suffix','nv_form')]

# cvc form or cc form
d3a$type <- 'v'
d3b$type <- 'nv'

# rename them
names(d3a) <- c('stem','category','exponent','base','suffix','word','type')
names(d3b) <- c('stem','category','exponent','base','suffix','word','type')

# make it long
d3 <- rbind(d3a,d3b)

# we don't want "ik", that's not a target set, it's a reference set
d3 <- droplevels( subset( d3, suffix != 'ik'))
# 93 stems, 2 * 5 Q forms per stem
```

Now we have 93 stems (111-18 defective stems). 

## 2. query forms on szoszablya web bulk search

We go to szotar.mokk.bme.hu/szoszablya/searchq.php and bulk search the forms.

```{r }
hits <- read.delim('webcorpus_query_results.txt', sep=';')
hits %>% head %>% kable
```

## 3. filter query for morphological analysis

```{r }
# get morphological parser tags

hits %>% group_by(analysis) %>% summarise(no_rows = length(analysis)) %>% arrange(no_rows) -> label_list
# eyeballing ensues

# pick the rare ones, since -- ideally -- we would see at least 93 of each tag if one tag covered one type of suffix (which it doesn't). this is because each Q suffix exists in at least CC form -- and we have 93 stems (sort of)
label_list %>% filter(no_rows < 10) -> suspicious_labels_list
kable(suspicious_labels_list)

# let's look at the forms themselves
hits %>% filter(analysis %in% suspicious_labels_list$analysis) -> suspicious_hits
# eyeballing ensues
kable(suspicious_hits)

# elhangzottak is out
hits %>% filter(!(word %in% c("elhangzottak"))) %>% droplevels() -> hits

# whatever this is is out
hits %>% filter(!(analysis %in% c("NOUN<POSS<2>><CAS<DAT>>","NOUN<PLUR>"))) %>% droplevels -> hits

# also this is utf 8 should be utf 16 but this is what is affected by the wrong coding
hits$word <- gsub('û', 'ű', hits$word)
hits$word <- gsub('õ','ő', hits$word)
hits$word <- tolower(hits$word)

# get hits with frequency (this is what's relevant)
hits %>% select(word, freq) -> hits1

# get lemmafreqs for each word
hits %>% select(word, lemmafreq) -> hits2
hits2$word <- as.character(hits2$word)
d2$word <- as.character(d2$word)

# we need to figure out the overall frequency of each stem. this is the sum of the unique lemma frequencies for the words that belong to the stem. you might want to read this line again.

# gather lemma freqs for stem
merge(hits2,d2, all = T) -> lemmaFreq 
# if a form is unattested, it's freq is zero.
lemmaFreq[is.na(lemmaFreq$lemmafreq),]$lemmafreq <- 0
# we need the stems and the various relevant lemma freqs
lemmaFreq %>% select(stem, lemmafreq) -> lemmaFreq
# we need to order this so we can sort it (I guess)
lemmaFreq[order(lemmaFreq$lemmafreq),] -> lemmaFreq
# for each stem
stemFreq <- ddply(lemmaFreq, "stem", function(x) {
# get unique values (those belong to the various lemmata assigned to this stem)
  unique <- unique(x$lemmafreq)
# sum these values  
  sumfreq <- sum(unique)
})

# name it stem freq
names(stemFreq)[2] <- 'stemfreq'

# we need the frequency of the -ik forms for each stem as well. pick forms
d %>% filter(v_suffix=='ik') %>% select(nv_form) -> ikforms
# pick freq for these forms from hit
hits1 %>% filter(word %in% ikforms$nv_form) %>% droplevels -> ikFreq
# get the biggest frequency (I'm lazy)
ikFreq[ order( ikFreq$word, -ikFreq$freq ), ] -> ikFreq
ikFreq %>% filter(!duplicated(word)) -> ikFreq
names(ikFreq) <- c('ik','ikfreq')
# pair up stems with ik forms
d %>% filter(nv_suffix=='ik') %>% select(stem,nv_form) -> stemIk
names(stemIk) <- c('stem','ik')
# pair up stems, ik forms, ik freq, stem freq
ikFreq <- merge(ikFreq,stemIk, all = T)
stems <- merge(ikFreq,stemFreq, all = T)

# add freqs to forms. get forms that are not 'ik'
hits3 <- hits1[!(grepl('ik$', hits1$word)),]
# pair em up with the long format target set
forms <- merge(d3, hits3, all = T)
# if it's not there its freq is zero
forms[is.na(forms$freq),]$freq <- 0

# i'ma save this
forms2 <- merge(forms, stems)
# write.csv(forms2, file='formatted_hits_long_current.csv', row.names = F)
```

## 4. make odds pairs

```{r }
# make odds pairs. make sure it's one pair per suffix (no "dohanyzhat" etc)

# these stems are beyond saving
forms %>% filter(!(stem %in% c("ér","fej","por","szag"))) -> forms

# get two types of forms
forms_a <- droplevels( subset( forms, type == 'v'))
forms_b <- droplevels( subset( forms, type == 'nv'))

# something ain't right. it should be same number of rows.

forms_a %>% mutate(index = paste(stem, exponent)) %>% arrange(index) %>% mutate(dupl = duplicated(index)) -> forms_a
forms_b %>% mutate(index = paste(stem, exponent)) %>% arrange(index) %>% mutate(dupl = duplicated(index)) -> forms_b

# some entries are repeated
forms_a %>% select(dupl) %>% summary
forms_b %>% select(dupl) %>% summary

# tOk and tAk merge in front stems. 
forms_a %>% filter(dupl==T) %>% kable
# I have a siking feeling it's all past pl tAk
forms_a %>% filter(dupl==T) %>% select(word) -> duplwords
hits %>% filter(word %in% duplwords$word) %>% select(word,analysis) %>% kable
# yup, they are -- at least according to the tagger -- with the exception of szagoltak. i'm removing it above from forms
# so within the duplicating subset of forms_a, the tOk forms are ghosts
forms_a %>% filter(dupl==F | exponent=='[tAk]') -> forms_a

# some OttAk forms appear twice. 
forms_b %>% filter(dupl==T) %>% kable
# I pick the one with a higher freq
forms_b %>% arrange(-freq) %>% filter(dupl==F) -> forms_b
# well forms_a still has two more forms than forms_b. will check that later*

# don't need these no more
forms_a$type <- NULL
forms_b$type <- NULL
forms_a$dupl <- NULL
forms_b$dupl <- NULL

names(forms_a) <- c("v_form","stem","category","exponent","v_base","v_suffix","v_freq","index")
names(forms_b) <- c("nv_form","stem","category","exponent","nv_base","nv_suffix","nv_freq","index")

pairs <- merge(forms_a, forms_b)

# something's off again. too many rows

pairs %>% mutate(dupl = duplicated(index)) -> pairs
pairs %>% filter(dupl==T) %>% select(index) -> duplforms
pairs %>% filter(index %in% duplforms$index) %>% kable
# right. i'll pick the more frequent forms.
pairs %>% arrange(-nv_freq) %>% filter(!duplicated(index)) -> pairs

# at least this line works fine
pairs <- merge(pairs, stems)
```

## Remove outliers (with 0 of either CVC or CC forms)

```{r }
var_counts <- ddply(pairs, 'stem', summarise, sum_v = sum(v_freq), sum_nv = sum(nv_freq))
var_counts$outlier <- ifelse(var_counts$sum_v < 1 | var_counts$sum_nv < 1, T, F)
var_counts %>% filter(outlier==F) %>% select(stem) -> var_stems
# var_counts %>% mutate(sum_all = sum_v + sum_nv) -> var_counts
# d %>% group_by(stem) %>% filter(v_freq > 0 | nv_freq > 0) %>% ddply(.(stem), n_q_types = nrow) -> d2
# d2 <- droplevels( subset(d, d$stem %in% var_stems$stem))
pairs %>% filter(stem %in% var_stems$stem) -> pairs
```

## 5. Add summary stats

```{r }
pairs$odds <- ( pairs$v_freq + 1 ) / ( pairs$nv_freq + 1) # Harald Baayen does it so it's fine
pairs$log_odds <- log(pairs$odds)

pairs$form_freq <- pairs$v_freq + pairs$nv_freq
pairs$log_form_freq <- log(pairs$form_freq + 1)

pairs$log_ikfreq <- log(pairs$ikfreq)
pairs$log_stemfreq <- log(pairs$stemfreq)

pairs <- pairs %>% group_by(ik) %>% mutate(stem_nv_freq = sum(nv_freq), stem_v_freq = sum(v_freq), stem_odds = (stem_v_freq + 1 ) / ( stem_nv_freq + 1), log_stem_odds = log(stem_odds))


plot(density(pairs$log_odds, bw = 0.4), xlim = c(-10,10), main = 'log odds for forms, all forms')
abline(v = 2.5*mad(pairs$log_odds), col = 'red', lty = 2)
abline(v = -2.5*mad(pairs$log_odds), col = 'red', lty = 2)
# the distribution is far less horrible in the first place, since it only contains -- more or less -- varying suffixes

# we either remove STEMS based on SUMMED log odds outside 2.5 median absolute deviation
# (this removes two forms)
# -- or, rather, flag stems WITH such forms

# flag outlier pairs
pairs$outlier_form <- ifelse(pairs$log_odds > 2.5*mad(pairs$log_odds) | pairs$log_odds < -2.5*mad(pairs$log_odds), T, F)
# flag stems with outlier pairs
pairs %>% group_by(stem) %>% summarise(outlier_forms = sum(outlier_form)) -> thing
thing$outlier_stem <- ifelse(thing$outlier_forms > 0, T, F)
thing %>% select(stem, outlier_stem) -> thing

pairs <- merge(pairs, thing)

plot(density(pairs[pairs$outlier_stem==F,]$log_odds, bw=0.4), xlim=c(-10,10), main='log odds for forms, without outliers')
abline(v = 2.5*mad(pairs$log_odds), col = 'red', lty = 2)
abline(v = -2.5*mad(pairs$log_odds), col = 'red', lty = 2)

# Now we create various metrics of number of Q suffixed forms/ frequency of ~ etc

# stem freq - ik freq

pairs$diff_freq <- pairs$stemfreq - pairs$ikfreq
pairs$log_diff_freq <- log(pairs$diff_freq)

# number of existing q forms

pairs %>% group_by(stem) %>% filter(v_freq > 0) %>% summarise(n_cvc_q_types = n()) -> var_sums1
pairs %>% group_by(stem) %>% filter(nv_freq > 0) %>% summarise(n_cc_q_types = n()) -> var_sums2
var_sums <- merge(var_sums1,var_sums2)
var_sums %>% mutate(n_q_types = n_cvc_q_types + n_cc_q_types) %>% select(stem,n_q_types) -> var_sums
pairs <- merge(pairs,var_sums, all=T)

# and some extra formatting.

# consonant cluster
pairs$cluster <- str_match(pairs$nv_base, "[aáeéiíoóöőuúüű].{2,3}$")
pairs$cluster <- sub('^.','', pairs$cluster)

pairs$cluster <- sub('^ng','g', pairs$cluster)
pairs$C1 <- substr(pairs$cluster, 1, nchar(pairs$cluster)-1)
pairs$C2 <- substr(pairs$cluster, nchar(pairs$cluster), nchar(pairs$cluster))

# features
labial <- c('p','b','m','f','v')
coronal <- c('t','d','n','j','l','r','sz','z')
palatal <- c('ty','gy','ny','s','zs')
velar <- c('k','g','h')

stop <- c('p','b','t','d','ty','gy','k','g')
fricative <- c('f','v','sz','z','s','zs','h')
nasal <- c('m','n','ny')

voiced <- c('b','d','gy','g','v','z','zs')
voiceless <- c('p','t','ty','k','f','sz','s')

# is cluster homorganic?

pairs$homorganic <- ifelse(
  pairs$C1 %in% labial & pairs$C2 %in% labial |
    pairs$C1 %in% coronal & pairs$C2 %in% coronal |
      pairs$C1 %in% palatal & pairs$C2 %in% palatal |
        pairs$C1 %in% velar & pairs$C2 %in% velar,
  T,
  F
) # this is how you don't code, kids

# View(unique(pairs[,c('C1','C2','homorganic')]))

# sonority scale
doSonority <- function(x){
x <- gsub("(j|ly)", 1, x)
x <- gsub("(r)", 2, x)
x <- gsub("(l)", 3, x)
x <- gsub("(m|n|ny)", 4, x)
x <- gsub("zs", 5, x)
x <- gsub("sz", 6, x)
x <- gsub("(v|z)", 5, x)
x <- gsub("(f|s)", 6, x)
x <- gsub("(b|d|g|gy)", 7, x)
x <- gsub("(p|t|k|ty)", 8, x)
return(x)
}

pairs$C1son <- doSonority(pairs$C1)
pairs$C2son <- doSonority(pairs$C2)
# hist(as.numeric(pairs$C1son))
# hist(as.numeric(pairs$C2son))
pairs$sonority <- as.numeric(pairs$C2son) - as.numeric(pairs$C1son)
pairs$syl_count <- str_count(pairs$stem, '[aáeéiíoóöőuúüű]')
```

We pull in the list of stems from Rebrus. These are annotated by him -- a Hungarian linguist. He also removed a number of forms. 

```{r }
rebrus <- read.delim('epenthetic-ik-verbs-features.csv', sep=',')
names(rebrus)[names(rebrus)=='lemma'] <- 'ik'
rebrus %>% select(ik,compound,free.stem,intransitive) -> rebrus
pairs <- merge(pairs,rebrus)
```

Rebrus excluded:

- feneklik (cf. fenekel)
- habzik (cf. haboz)
- nyálzik (cf. nyálaz)
- sugárzik (cf. sugároz)
- torlik (cf. torol)
- vedlik (cf. vedel)

We add glosses and phonetic transcription for plots.

```{r }
stem_matcher <- pairs %>% select(ik) %>% unique

gloss_stem <- c('flow','decompose','toddle','burgeon','stink','perish','smoke','be voiced','peel','tooth','choke','bend','die','sound','be missing','wave','light up','fog','breathe','get goosebumps','collapse','pour','decompose','occur','surge','mate','efflux','swarm','degrade','rally','voice','tantrum','shine','jump','bleed','storm','lighten','bloom','relate')
phon_stem <- c("aːrɒmlik", "bomlik", "botlik", "burjaːnzik", "byːzlik", "døglik", "dohaːɲzik", "ɛlhɒngzik", "fɛʃlik", "fogzik", "fuldoklik", "hɒjlik", "hɒldoklik", "hɒngzik", "hiaːɲzik", "hullaːmzik", "kivilaːglik", "kødlik", "leːlɛgzik", "lúdbøːrzik", "omlik", "ømlik", "oslik", "øtlik", "øzønlik", "paːrzik", "pɒtɒkzik", "rɒjzik", "romlik", "ʃɛrɛglik", "soːlɒmlik", "tɒjteːkzik", "tyndøklik", "ugrik", "veːrzik", "vihɒrzik", "vilaːglik", "viraːgzik", "visoɲlik")

stem_matcher <- cbind(stem_matcher,gloss_stem,phon_stem)

pairs <- merge(pairs,stem_matcher)

```

```{r }
# ...profit

write.csv(pairs, file='formatted_filtered_hits_pairs_current.txt', row.names = F)
```
